Describing demographic models is hard

### 2020-05-15 paper.tex
martin bug

To illustrate the demography API, \msprime\ included a description of the
three population Out-of-Africa model~\citep{gutenkunst2009inferring}
as part of its tutorial documentation. This example model, however,
was not correctly implemented.

@@

[Description of what went wrong in the AJHG paper~\citep{martin2017human}.
Some analysis of how this affects the conclusions of the paper.]

[How was the error found? I wonder if the reason that Alicia didn't
do any basic checks on the simulation output was because it was
so massive. Until recently we couldn't do things like compute Fst, etc
on the tree sequences. Still, good enough reason to run small simulations
first to make sure that you're doing something sensible.]

This error was clearly caused by poor API design. By forcing the user to
pass around three separate parameters to describe the demographic model,
it's inviting simple mistakes such as this. In \msprime\ 1.0 we introduce
a Demography class, which wraps these three parameters.
Thus, instread of writing
\begin{lstlisting}[frame=single]
dbg = msprime.DemographyDebugger(
    population_configurations=population_configurations,
    migration_matrix=migration_matrix,
    demographic_events=demographic_events)
dbg.print_history()
ts = msprime.simulate(
    population_configurations=population_configurations,
    migration_matrix=migration_matrix,
    demographic_events=demographic_events)
\end{lstlisting}
we would now write
\begin{lstlisting}[frame=single]
dbg = demography.debug()
dbg.print_history()
ts = msprime.simulate(demography=demography)
\end{lstlisting}

>>>>

To illustrate the demography API, \msprime\ included a description of a widely-used
three population Out-of-Africa model~\citep{gutenkunst2009inferring}
as part of its tutorial documentation. This example model, however,
was not correctly implemented.

@@

In another publication using this model~\citep{martin2017human}, 
another error crept in: the model itself was defined as suggested 
in the documentation, and tested using model visualization and 
debugging tools. After these checks were performed, however, the simulation 
itself was finally performed with an incorrect parameter, leading to a vast 
overestimate of the divergence across human populations.

In this case the error had a large effect on the simulation output and 
the conclusions of the analysis. This simulation was performed to assess 
the transferabilty of polygenic risk scores across human populations.

Unfortunately, whereas the correct model suggests a mean $F_{ST}$ of 
XXX across the three population, the simulated model generated an $F_{ST}$ of 
XXX, exaggerating the difficulty in transferring genetic risk across populations (the difficulty still exists, Figure X).  The resulting
publication has been influential in the discussion of health inequalities and 
genomics, [[with over 350 citations since 2017 ?]].

This particular error suggests three lessons.

First, the msprime API design was poorly designed, requiring the user to pass
three separate parameters to specify a demographic model. 
 In \msprime\ 1.0 we introduce
a Demography class, which wraps these three parameters.
Thus, instread of writing
\begin{lstlisting}[frame=single]
dbg = msprime.DemographyDebugger(
    population_configurations=population_configurations,
    migration_matrix=migration_matrix,
    demographic_events=demographic_events)
dbg.print_history()
ts = msprime.simulate(
    population_configurations=population_configurations,
    migration_matrix=migration_matrix,
    demographic_events=demographic_events)
\end{lstlisting}
we would now write
\begin{lstlisting}[frame=single]
dbg = demography.debug()
dbg.print_history()
ts = msprime.simulate(demography=demography)
\end{lstlisting}.

Second, testing of the final simulated data is important. This can be challenging, 
because the amount of simulated data can be large. However, recent progress in 
computing summary statistics from tree sequence data \cite{} can make this easier. 
Even if the entire simulation data cannot be easily evaluated, subsets of the data 
can be examined to identify large error.

Third, open data analysis pipelines are necessary for the self-correcting nature of science. 
This subtle error of large effect was only discovered as one of us was trying to take
advantage of the simulation pipeline developed in~\citep{martin2017human} to pursue 
additional analyses on a similar topic. We identified the bug by tracing down an unexpected 
result in our analysis.
